### 2023-05-26 src/analysis/Analysis.Rmd

---
title: 'Analyses'
author: 
output: 
  github_document:
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r "setup", include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center",
                      fig.path=('../../report/figures/'))
library(rmarkdown)
```

# Required Packages &amp; Reproducibility
```{r "env", message=FALSE, warning=F}
rm(list=ls())
source(here::here("src/lib/functions.R"))
```

# Analyses
```{r "data for analyses"}
load(here("data/intermediate/cleaned_data_nl.RData"))
load(here("data/intermediate/cleaned_data_us.RData"))
```

## Balance Checks

The figure below shows that the data is not unbalanced for any of the variables: As described in the Pre-Analysis Plan (p.10), we will not add any covariates to the analyses as controls.

```{r "balance-checks", fig.height=12, fig.width=8}
source(here("src/analysis/balance-test.R"))
df
```

## Baseline Model
```{r "baseline", fig.align="center", fig.height=4, fig.width=8}
source(here("src/analysis/baseline.R"))
b

b_b
```

```{r "issues", fig.align="center", fig.height=16, fig.width=8}
source(here("src/analysis/issue_specific.R"))
b_i

b_ib
```

## Ideological Bias hypothesis
```{r "h1", fig.align="center", fig.height=8, fig.width=8}
source(here("src/analysis/h1.R"))
p1

p1b

p1 / p1b +
  plot_layout(heights = c(1, 1.3)) + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom')

p1_2 / p1b2 +
  plot_layout(heights = c(1, 1.3)) 
```

## Overinterpretation Hypotheses
```{r "h2", fig.align="center", fig.height=8, fig.width=8}
source(here("src/analysis/h2.R"))
p2

p2b

p2c
```

## Masking Hypothesis
```{r "h3", fig.align="center", fig.height=8, fig.width=8}
source(here("src/analysis/h3.R"))
p3

p3b

p3c
```

## Exploration
```{r "exploration", fig.height=7, fig.width=8}
source(here("src/analysis/exploration.R"))
p_e1
p_e2
p_e3_1
p_e3_2
p_b
p_b2

source(here("src/analysis/profile_resp.R"))
kbl(e1, booktabs =T, caption = "\\label{tab:conditions}Profile Dutch Stance Annotators") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "7cm")

kbl(e1_us, booktabs =T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "7cm")
```