### 2023-05-26 src/data-processing/prepare-data.Rmd

---
title: 'Prepare Data'
author: 
output: 
  github_document:
    toc: yes
always_allow_html: true
editor_options:

  chunk_output_type: console
---

```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center",
                      fig.path=('../../report/figures/'))
library(printr)
options(digits = 2)
library(rmarkdown)
```

## Required Packages &amp; Reproducibility
```{r "env", message=FALSE, warning=F}
rm(list=ls())
source(here::here("src/lib/functions.R"))
#renv::snapshot()
```

## Tidy Data
This code chuck downloads the data from Qualtrics using the API and cleans the raw data.

```{r "get & clean data", results="hide"}
#Clean Dutch Data
d_nl <- read_sav(here("data/raw-private/NLOZ2206_Mariken_Bias-experiment.sav"))
source(here("src/data-processing/clean_data_nl.R"))

#Clean US Data
d_us <- fetch_survey(surveyID = "SV_86XBcsFUJka9eke", 
                    verbose = TRUE, force_request = T,
                    label = FALSE, convert = FALSE)
source(here("src/data-processing/clean_data_us.R"))
```

## Save Data for Analysis
```{r "save data"}
save(d_nl, file = here("data/intermediate/cleaned_data_nl.RData"))
save(d_us, file = here("data/intermediate/cleaned_data_us.RData"))
```

## Visualization of Data

### Dependent Variable by Issue
```{r "Dependent Variable", echo=F, fig.width=15, fig.height=8}
p1 <- d_nl %>%
  group_by(stance,issue,specification, masking) %>% 
  count() %>% 
  mutate(stance = ifelse(stance==0, "Not Correctly identified stance",
                         "Correctly identified stance"),
         specification = ifelse(specification == "Specified", 
                                "Condition: Specified Sentence", "Underspecified Sentence"),
         masking = ifelse(masking == "Masked", "Actor Masked", "Party Revealed")) %>% 
  ggplot(aes(y = issue, x = n, fill = stance)) +
  geom_col() +
  facet_grid(masking~specification) +
  labs(x = "", y = "", 
       title = "Dutch Data") +
  #scale_x_continuous(labels = scales::percent) +
  theme_ipsum() +
  scale_fill_manual(values = fig_cols[c(3, 2)]) +
  theme(legend.position = "bottom",
        legend.title = element_blank())

p2 <- d_us %>%
  group_by(stance,issue,specification, masking) %>% 
  count() %>% 
  mutate(stance = ifelse(stance==0, "Not Correctly identified stance",
                         "Correctly identified stance"),
         specification = ifelse(specification == "Specified", 
                                "Condition: Specified Sentence", "Underspecified Sentence"),
         masking = ifelse(masking == "Masked", "Actor Masked", "Party Revealed")) %>% 
  ggplot(aes(y = issue, x = n, fill = stance)) +
  geom_col() +
  facet_grid(masking~specification) +
  labs(x = "", y = "", 
       title = "US Data") +
  #scale_x_continuous(labels = scales::percent) +
  theme_ipsum() +
  scale_fill_manual(values = fig_cols[c(3, 2)]) +
  theme(legend.position = "bottom",
        legend.title = element_blank())

p1 + p2 + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom')

```

### Covariates
```{r "Independent Variables", echo=F, message=FALSE, warning=FALSE,fig.width=10, fig.height=10}
rm(p1,p2)
p <- d_nl %>% select(distance, PT7, PT6) %>% 
  pivot_longer(cols = everything(),
               values_to = "score") %>% 
  group_by(name) %>% 
  summarise(m = mean(score),
            stdev = sd(score)) %>% 
  mutate(name = recode(name,
                       `distance` = "Ideological Distance \n(> = More Distance)",
                       `PT6` = "Ideological Self Placement \n(0 = Left, 10 = Right)",
                       `PT7` = "Political Knowledge \n(0 = Low, 6 = High)"),
         country = "The Netherlands")

p2 <- d_us %>% 
  select(distance, PT7, PT7b, PT6) %>% 
  pivot_longer(cols = everything(),
               values_to = "score") %>% 
  group_by(name) %>% 
  summarise(m = mean(score),
            stdev = sd(score)) %>% 
  mutate(name = recode(name,
                       `distance` = "Ideological Distance \n(> = More Distance)",
                       `PT6` = "Ideological Self Placement \n(0 = Extremely Liberal, 6 = Extremely Conservative)",
                       `PT7` = "Political Knowledge \n(0 = Low, 3 = High)",
                       `PT7b` = "Bullshit Receptivity \n(0 = Low, 5 = High)"),
         country = "United States")

p %>% 
  add_case(p2) %>% 
  ggplot(aes(y = name, x = m, 
             xmin = (m - (1.96*stdev)),
             xmax = (m + (1.96*stdev)),
             color = country)) +
    geom_point(position = position_dodge(.2)) +
    geom_errorbar(width = 0, position = position_dodge(.2)) +
    theme_ipsum() +
    scale_colour_manual(values = fig_cols) +
    labs(x = "Value", y = "") +
  theme(legend.position = "bottom",
        legend.title = element_blank())
```

### Correlations Matrix
```{r "Correlations Matrix", echo=F, message=F, warning=F, fig.width=15, fig.height=15}
d_nl %>%
  drop_na() %>% 
  select(`Age` = D1,
         `Gender` = D2,
         `Education` = D3,
         `Region`  = D4,
         `Income`  = D5,
         `Job`  = D6,
         `Vote Recall`  = PT1,
         `Position: Immigration`  = PT2,
         `Position: Climate`  = PT3,
         `Position: Tax`  = PT4,
         `Position: EU`  = PT5,
         `Ideology`  = PT6,
         `Political Knowledge` = PT7,
         `Distance to Political Party` = distance,
         `Stance` = stance) %>%
  ggcorrmat(
  type = "robust", # correlation method
  cor.vars = c(everything()), # a range of variables can be selected
  colors = c(fig_cols[1], fig_cols[4], fig_cols[8]),
  title = "Correlalogram for Variables under Study",
  subtitle = "Dutch Data"
)

d_us %>%
  drop_na() %>% 
  select(`Age` = D1,
         `Gender` = D2,
         `Education` = D3,
         `Region`  = D4,
         `Income`  = D5,
         `Job`  = D6,
         `PID`  = PT1,
         `PID Strength` = PT1b,
         `Position: Immigration`  = PT2,
         `Position: Climate`  = PT3,
         `Position: Tax`  = PT4,
         `Position: EU`  = PT5,
         `Ideology`  = PT6,
         `Political Knowledge` = PT7,
         `Bullshit Receptivity` = PT7b,
         `Distance to Political Party` = distance,
         `Stance` = stance) %>%
  ggcorrmat(
  type = "robust", # correlation method
  cor.vars = c(everything()), # a range of variables can be selected
  colors = c(fig_cols[1], fig_cols[4], fig_cols[8]),
  title = "Correlalogram for Variables under Study",
  subtitle = "US Data"
)
```